---
title: "Week 9, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rstanarm)
library(tidyverse)

# We will use the `kenya` dataset from PPBDS.data. This dataset is from Harris,
# Kamindo and Van der Windt (2020): "Electoral Administration in Fledgling
# Democracies: Experimental Evidence from Kenya." The authors worked with
# Kenya's electoral commission in 1,674 communities by assigning polling
# stations to either a control group or to one of five other treatments.

# For simplicity we will only consider polling stations assigned to either
# "control" or to "local". The latter means a treatment in which election
# officials visited the community's polling station for two days, thereby making
# it easier for people to register.

# The outcome variable, `reg_chg`, is the change in voter registration in each
# community. `poverty` is the percentage of the community below the poverty
# level. `distance` is the number of kilometers between the polling station and
# the central political office (where registrations normally take place).
# `pop_density` is a measure of population density around the polling station.


week_9 <- kenya %>% 
  rename(reg_chg = reg_byrv13) %>% 
  filter(treatment %in% c("control", "local")) %>% 
  droplevels() %>% 
  select(reg_chg, treatment, poverty, 
         distance, pop_density)
week_9
```


## Scene 1

**Prompt:** In addition to the outcome variable and the treatment, we have 3 statistics for the communities in the sample --- poverty, distance and population density. See code comments above for details. Never hurts to do a little exploratory data analysis. Look at the data!

* Make a scatter plot of `poverty` on the y-axis and `distance` in the x-axis. Include a `geom_smooth()` with `method` set to "lm" and `se` equal to FALSE. Write a sentence to answer these questions:

  + Does poverty cause distance? Or does distance cause poverty? Or both? Or neither?
  + How accurate is the model beyond 90 kilometers?  
  + Eye-balling the chart, how different is the poverty rate between two polling stations whose distance differs by 50 kilometers?
  
* Create a linear model which relates poverty to distance. This is similar to the models we created last week which related age to party. See the discussion in Chapter 9 for advice.  Write down the math. 
  
* Fit a `stan_glm()` model which relates poverty to distance. Print a summary. (Hint: In order to see the coefficient of `distance`, you will probably need to use the `digits` argument to `print()`.) Interpret the meaning of the two important parameters.  

```{r Q1}

# This is a predictive model, where both poverty and distance are variables that
# have not been "treated" or "controlled" by the experiment. This means that 
# no one variable here causes the other, even though there is a seemingly positive 
# correlation between greater distance from a polling station and poverty rate.

# There is only one observation beyond the 90 km line, so the model is not 
# very accurate. The whole linear model is pulled out by this observation.

# Along the linear model line, the difference in poverty rates between a place
# that is 50 km further from a poll than another is about 8 percentage points 
# (0.08), or the slope of the model.

ggplot(week_9, aes(distance, poverty)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# The linear model relating poverty to distance:
# poverty = intercept + coefficient * distance + sigma

stan_glm(formula = poverty ~ distance,
         data = week_9,
         refresh = 0) %>%
print(digits = 4)

# The Intercept parameter shows that the median of this dataset has a y-intercept
# of around 40%. This means that when the location is 0 km away from a polling
# station (it has a polling station right next to it), the median poverty rate
# is already starting at around 40%.
# The second parameter is supposed to show the median effect as distance increases
# by 1 km on poverty rates, but it is very small at a median of 0.18%. This means
# that the distance has a very small positive effect on poverty.

```
$$ poverty_i = \beta_0 + \beta_1 distance_{i} + \epsilon_i$$


## Scene 2

**Prompt:** Let's make a model with `reg_chg` as the outcome variable and with treatment, poverty, distance and pop_density as the predictors. (Using outcome/predictors is the typical nomenclature in statistics. In Economics, we would refer to dependent/independent variables.)

* Provide the math for this model.

* Estimate the model. Assign the fitted model to `fit_2`. Interpret the important parameters of the model.

```{r}

fit_2 <- stan_glm(formula = reg_chg ~ treatment + poverty + distance + pop_density,
         data = week_9,
         refresh = 0) %>%
  print(digits = 6)

# The intercept is the change in registration when all of the other variables/
# parameters are set to 0. This means treatment is control, poverty rate is 0, 
# distance from a polling station is 0 km, and population density is relatively
# 0, very low density? This median value is a small negative number, - 3.4 %. 
# Registration drops when these variables are all set to 0.

# On median, treating people has a more positive effect on change in registration
# than the (negative) effect of controlling this variable.
# Poverty shows a slight positive slope for changes in registration as the 
# poverty rate increases one unit.
# Distance shows a slight negative slope for changes in registration as the 
# distance from a polling station increases one km.
# pop_density shows a larger negative slope for changes in registration
# as the population density increases.

```

$$ regchange_i = \beta_0 + \beta_1 treatment_{i} + \beta_2 poverty_{i} + \beta_3 distance_{i} + \beta_4 popdensity_{i} + \epsilon_i$$

